---
title: "MAS5312_Statistical_Machine_Learning_Assignment_01"
author: "Group_1"
output: word_document
---

## Prerequisits

```{r}
# Prerequisits
library(ggplot2)
library(sf)
library(geojsonR)
library(ggpubr)
library(reshape2)
library(ggcorrplot)
library(tidyverse)
library(dplyr)
setwd("C:/Personal/Books/MSc. in Applied Statistics/Semester_3/MAS5312_Statistical_Machine_Learning/Dataset")
```

## Load Data

```{r }
# Load and preprocess data
df <- read_csv('Dataset/train.csv')
df <- df %>%  mutate(
  id = as.character(id),
  Driving_License = ifelse(Driving_License == 1,"Licensed","Not Licensed"),  
  Region_Code = as.character(Region_Code),
  Previously_Insured = ifelse(Previously_Insured == 1,"Insured","Not Insured"),  
  Vehicle_Damage = ifelse(Vehicle_Damage == "Yes", "Damaged", "Not Damaged"),
  Policy_Sales_Channel = as.character(Policy_Sales_Channel),
  Response = ifelse(Response == 1, "Interested", "Not Interested")) 
```

```{r}
summary(df)
```
Change data types
```{r}
# df <- df %>%  mutate(
#   id = as.character(id),
#   Driving_License = ifelse(Driving_License == 1,"Licensed","Not Licensed"),  
#   Region_Code = as.character(Region_Code),
#   Previously_Insured = ifelse(Previously_Insured == 1,"Insured","Not Insured"),  
#   Vehicle_Damage = ifelse(Vehicle_Damage == "Yes", "Damaged", "Not Damaged"),
#   Policy_Sales_Channel = as.character(Policy_Sales_Channel),
#   Response = ifelse(Response == 1, "Interested", "Not Interested")) 
```



## Gender wise analysis

Gender proportions

```{r}
df %>% group_by(Gender) %>% summarise(cnt = n()) %>% mutate(percent = cnt/sum(cnt))
```
```{r} 
df %>% group_by(Gender, Response) %>% summarise(cnt = n()) %>% mutate(percent = cnt/sum(cnt))
```


## Age analysis

```{r}
# Figure 1: Gender wise age distribution
df %>% select(Gender,Age) %>% arrange(Gender) %>% 
  ggplot(aes(Age)) +
  geom_boxplot(mapping = aes(x = Age, y = Gender , fill = Gender))+
  labs(title = "Gender wise age distributions",
       x = "Age",
       y = NULL) +
  theme_minimal() +
  theme(legend.position="none")
```

```{r}
# Gender wise age quartile margins
df %>% filter(Gender=='Male') %>% select(Age) %>% quantile(na.rm = TRUE)
df %>% filter(Gender=='Female') %>% select(Age) %>% quantile(na.rm = TRUE)
```
```{r}
#df[1:10]
```
```{r}
# Define gender wise age quartile categories
df_age <- df %>% select('Gender', 'Age', 'Response')
df_age$Age_Quartile <- cut(
  df_age$Age,
  quantile(df_age$Age),include.lowest=TRUE,labels= c('1','2','3','4'))
```


```{r}
df_age %>% group_by(Gender, Age_Quartile) %>% summarise(cnt = n()) %>% mutate(percent = cnt/sum(cnt))
```
```{r}
df_age
```


```{r}
df_age_m <- df %>% filter(Gender=='Male') %>% select('Gender', 'Age', 'Response')
df_age_m$Age_Quartile <- cut(df_age_m$Age,quantile(df_age_m$Age),include.lowest=TRUE,labels= c('1','2','3','4'))

df_age_f <- df %>% filter(Gender=='Female') %>% select('Gender', 'Age', 'Response')
df_age_f$Age_Quartile <- cut(df_age_f$Age,quantile(df_age_f$Age),include.lowest=TRUE,labels= c('1','2','3','4'))

df_age = rbind(df_age_m,df_age_f)

df_age %>% group_by(Gender, Age_Quartile, Response) %>% summarise(cnt = n()) %>% mutate(percent = cnt/sum(cnt)*100) %>% 
  ggplot(aes(x=Age_Quartile, y=percent, fill=Response)) +
  geom_bar(position="dodge", stat="identity") + facet_wrap(~Gender, ncol=2) +
  geom_text(aes(label = paste(round(percent,2), "%")),vjust = -.5) +
  labs(title = "Age quartile wise male and female responses", x = "Age Quartile") +
  theme_minimal()+
  theme(
      strip.text = element_text(face = 'bold'),
      plot.caption = element_text(face = 'italic'),
      panel.grid.major = element_line('white', size = 0.5),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      axis.text.y=element_blank(),
      axis.ticks.y=element_blank()
    )
```
```{r}
df_age %>% filter(Response=="Interested")%>% group_by(Gender, Age_Quartile) %>% summarise(cnt = n()) %>% mutate(percent = cnt/sum(cnt)*100) %>% 
  ggplot(aes(x=Age_Quartile, y=percent, fill=Gender)) +
  geom_bar(position="dodge", stat="identity") + facet_wrap(~Gender, ncol=2) +
  geom_text(aes(label = paste(round(percent,2), "%")),vjust = -.5) +
  labs(title = "Positive responses - male and female age quartile percentages", x = "Age Quartile") +
  theme_minimal()+
  theme(
      strip.text = element_text(face = 'bold'),
      plot.caption = element_text(face = 'italic'),
      panel.grid.major = element_line('white', size = 0.5),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      axis.text.y=element_blank(),
      axis.ticks.y=element_blank()
    )
```



## Area wise analysis
```{r}
d1 <- df %>% group_by(Region_Code, Response) %>% summarise(cnt = n()) %>%
  mutate(percent = cnt/sum(cnt)*100) %>% arrange(Region_Code) 
d1 <- d1 %>%  mutate(order = as.numeric(Region_Code))

d2 <- d1 %>% filter(Response=='Not Interested') %>% select(percent)

d1 %>% ggplot(aes(x= reorder(Region_Code, order), y=percent, fill=Response)) +
  geom_bar(stat="identity")+
  labs(title = "Region wise Response Percentages", y = "%", x="Region Code") +
  theme(
      strip.text = element_text(face = 'bold'),
      plot.caption = element_text(face = 'italic'),
      panel.grid.major = element_line('white', size = 0.5),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
    )+
  geom_hline(aes(yintercept=mean(d2$percent))) + 
  annotate("text", x=50, y=round(mean(d2$percent),2)-2,
           label=paste(round(mean(d2$percent),2),"%"))+
  geom_hline(yintercept=mean(d2$percent)-2*sd(d2$percent), linetype='dotted') +
  annotate("text", x=50, y=mean(d2$percent)-2*sd(d2$percent), label="-2*sd") +
  geom_hline(yintercept=mean(d2$percent)+2*sd(d2$percent), linetype='dotted') +
  annotate("text", x=50, y=mean(d2$percent)+2*sd(d2$percent), label="+2*sd")

ggsave("region_wise_response.jpg", width = 30, height = 10, units = "cm")
```
```{r}
print(sd(d2$percent))
```



## Insurance impact on Response
```{r}
df %>% group_by(Previously_Insured, Response) %>% summarise(cnt = n()) %>% 
  mutate(percent = cnt/sum(cnt)*100)
```


```{r}
df %>% group_by(Previously_Insured, Response) %>% summarise(cnt = n()) %>% 
  mutate(percent = cnt/sum(cnt)*100) %>% 
  ggplot(aes(x=Response, y=percent, fill=Response)) +
  geom_bar(position="dodge", stat="identity") + 
  facet_wrap(~Previously_Insured, ncol=2) +
  geom_text(aes(label = paste(round(percent,2), "%")),vjust = -.5) +
  labs(title = "Previously Insured Status effect on Response", y = NULL) +
  theme_minimal()+
  theme(
      strip.text = element_text(face = 'bold'),
      plot.caption = element_text(face = 'italic'),
      panel.grid.major = element_line('white', size = 0.5),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      axis.text.y=element_blank(),
      axis.ticks.y=element_blank()
    )
```
## Figure 6: Driving license status effect on the response
```{r}
df %>% group_by(Driving_License) %>% summarise(avg_age = mean(Age))

df %>% group_by(Driving_License, Response) %>% summarise(cnt = n()) %>% 
  mutate(percent = cnt/sum(cnt)*100) %>% 
  ggplot(aes(x=Response, y=percent, fill=Response)) +
  geom_bar(position="dodge", stat="identity") + 
  facet_wrap(~Driving_License, ncol=2) +
  geom_text(aes(label = paste(round(percent,2), "%")),vjust = -.5) +
  labs(title = "Driving License status effect on Response", y = NULL) +
  theme_minimal()+
  theme(
      strip.text = element_text(face = 'bold'),
      plot.caption = element_text(face = 'italic'),
      panel.grid.major = element_line('white', size = 0.5),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      axis.text.y=element_blank(),
      axis.ticks.y=element_blank()
    )
```
## Vehicle Age impact on Premium and Response
```{r}
df %>% group_by(Vehicle_Age) %>% summarise(avg_premium = mean(Annual_Premium))

d3 <- df %>% group_by(Vehicle_Age, Response) %>% summarise(cnt = n()) %>%
  mutate(percent = cnt/sum(cnt)*100)
d3 <- d3 %>%  mutate(age_order = case_when(Vehicle_Age=="< 1 Year"~1,
                                             Vehicle_Age=="1-2 Year"~2,
                                             TRUE ~ 3))

d3 %>% ggplot(aes(x=reorder(Vehicle_Age, age_order), y=percent, fill=Response))+
  geom_bar(position="dodge", stat="identity") +
  geom_text(aes(label = paste(round(percent,2), "%")),vjust = -.5) +
  labs(title = "Vehicle Age impact on Premium and Response", x="Vehicle Age") +
  theme_minimal()+
  theme(
      strip.text = element_text(face = 'bold'),
      plot.caption = element_text(face = 'italic'),
      panel.grid.major = element_line('white', size = 0.5),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      axis.text.y=element_blank(),
      axis.ticks.y=element_blank()
    )
```


## Vehicle Damage impact on Premium and Response
```{r}
df %>% group_by(Vehicle_Damage) %>% 
  summarise(Average_Premium = mean(Annual_Premium))

df %>% group_by(Vehicle_Damage, Response) %>% summarise(cnt = n()) %>%
  mutate(percent = cnt/sum(cnt)*100) %>% 
  ggplot(aes(x=Vehicle_Damage, y=percent, fill=Response)) +
  geom_bar(position="dodge", stat="identity") + 
  #scale_x_discrete(labels=c("< 1 Year", "1-2 Year", "> 2 Years")) +
  geom_text(aes(label = paste(round(percent,2), "%")),vjust = -.5) +
  labs(title = "Damage Status impact on Premium and Response", 
       x="Vehicle Condition") +
  theme_minimal()+
  theme(
      strip.text = element_text(face = 'bold'),
      plot.caption = element_text(face = 'italic'),
      panel.grid.major = element_line('white', size = 0.5),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      axis.text.y=element_blank(),
      axis.ticks.y=element_blank()
    )
```


## Driving License impact on Premium and Response
```{r}
df %>% group_by(Driving_License) %>% 
  summarise(Average_Premium = mean(Annual_Premium))

df %>% group_by(Driving_License, Response) %>% summarise(cnt = n()) %>% 
  mutate(percent = cnt/sum(cnt)*100) %>%
  ggplot(aes(x=Driving_License, y=percent, fill=Response)) +
  geom_bar(position="dodge", stat="identity") +
  geom_text(aes(label = paste(round(percent,2), "%")),vjust = -.5) +
  labs(title = "Driving License impact on Premium and Response", 
       x="Driving License Status") +
  theme_minimal()+
  theme(
      strip.text = element_text(face = 'bold'),
      plot.caption = element_text(face = 'italic'),
      panel.grid.major = element_line('white', size = 0.5),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      axis.text.y=element_blank(),
      axis.ticks.y=element_blank()
    )
```
## Sales Channel impact on Response
```{r}
d4 <- df %>% group_by(Policy_Sales_Channel, Response) %>% 
  summarise(cnt = n()) %>% mutate(percent = cnt/sum(cnt)*100) %>% 
  arrange(percent) 

df %>% select(Policy_Sales_Channel) %>% unique() %>% count()

df %>% group_by(Policy_Sales_Channel) %>% summarise(cnt = n()) %>% 
  filter((cnt>=100)) %>% count()

d4 %>% filter((Response=="Interested") & (cnt>=100) & (percent>=15)) %>% 
  select(Policy_Sales_Channel)  %>% unique() %>% count()

d4 %>% filter((Response=="Interested") & (cnt>=100)) %>% 
  arrange(desc(percent)) %>% 
  ggplot(aes(x= reorder(Policy_Sales_Channel, percent), 
             y=percent, fill=Response)) +
  geom_bar(stat="identity")+
  coord_flip()+
  labs(
    title ="Sales channels with higher positive responses(min 100 responses)", 
    y = "%", x="Sales Channel") +
  theme_minimal()+
  theme(
      strip.text = element_text(face = 'bold'),
      plot.caption = element_text(face = 'italic'),
      panel.grid.major = element_line('white', size = 0.5),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
    )
```
```{r}
df %>% select(Response,Vintage) %>% arrange(Response) %>% 
  ggplot(aes(Vintage)) +
  geom_boxplot(mapping = aes(x = Vintage, y = Response , fill = Response))+
  labs(title = "Vintage Distribution",
       x = "Vintage",
       y = NULL) +
  theme_minimal() +
  theme(legend.position="none")
```



